Constructing Statistical Deformable Models with Shape Flow

### 2015-07-17 contents/conclusion.tex
Final Rebuttal

We proposed the first, to the best of our knowledge, methodology that can build a dense deformable model "in-the-wild". This methodology does not require a consistent set of landmarks, avoiding the tedious task of careful manual annotations. We have exploited state-of-the-art multiframe optical flow methodologies to establish accurate and robust dense correspondences among shapes, when represented in a landmarks-free fashion. The experimental evaluations and comparisons revealed that dense AAMs that are built on top of our methodology achieve very good performance and can work effectively even in cases of objects that cannot have consistent landmarks.

>>>>

In this paper, we propose the first, to the best of our knowledge, methodology for building dense deformable models "in-the-wild". Our approach does not require objects to be annotated with consistent sets of landmarks and can make use of highly-efficient (in terms of required annotation time) curve line annotations, easing the tedious task of careful manual annotation. We have exploited state-of-the-art multiframe optical flow techniques to establish accurate and robust dense correspondences among shapes, when represented in a landmarks-free fashion via the use of the SVS representation. The experimental evaluation and comparison revealed that dense AAMs that are built on top of our methodology achieve good performance and can work effectively even for objects for which is difficult to define a consistent set of landmarks (e.g. human ears).

### 2015-07-17 contents/evaluation.tex
Final Rebuttal

The Cumulative Error Distribution (CED) for this experiment is shown in Figure \ref{fig:svs_ced} while fitting statistics are reported in Table \ref{tab:svs_stats}. Looking at the results it can be clearly seen that using SVS results in a constant $\approx$5\% improvement over the simple binary shape representation for almost all CED graph and on the reported fitting statistics.

@@

\subsection{Constrained Basis}
\label{exp:basis}

This experiment investigate the impact of introducing subspace constrains for shape flow. The shape representations of training data are independence, thus simply applying flows on shapes breaks the fundamental assumption of motion smoothness. The hypothesis is able to be maintained for shapes as described in section~\ref{sec:trabasis}. To be specific, we take measures of point-to-point error for fitting same data with our pipeline with same configuration, but only alter the low rank basis.

The experiment results are manifested by constructing two dense AAMs, Basis-dAAMs and no-Basis-dAAMs,among 811 training images from Labelled Faces Parts in-the-Wild (LFPW)\cite{Belhumeur2011} database with 224 testing images using same annotation scheme by iBUG\footnoteref{ibug_300}. 
Cumulative error distributions are shown in Figure~\ref{fig:basis_ced} and corresponding statistics are listed in Table~\ref{tab:basis_stats}. 
It shows shape flow with constrained basis outperforms, constantly, standard flow technique by an unavoidable amount, which significantly supported the necessities of our low rank basis introduced in~\ref{sec:trabasis}.

@@

In the following set of experiments we use the proposed methodology to build dense AAMs of various object classes (faces, ears, bottles and body pose) and quantitatively assess their respective fitting accuracy when fitted to novel in-the-wild images of the previous objects.

>>>>

The Cumulative Error Distribution (CED) for this experiment is shown in Figure \ref{fig:svs_ced} while fitting statistics are reported in Table \ref{tab:svs_stats}. Looking at the results it can be clearly seen that using SVS results in an almost constant $\approx$10\% improvement over the simple binary shape representation for most of the CED graph and reported fitting statistics.

@@

\subsection{Correspondence Bases}
\label{exp:basis}
This experiment investigates the importance of applying subspace constraints to the shape flow estimation stage. To this end, we build and compare the performance of two different dense AAMs. The first one (SVS-AAMs) is built from deformation fields obtained by minimizing the variational shape flow energy proposed in section \ref{varshapeflow}, \ie constraining the shape flow with the correspondence bases described in section \ref{sec:trabasis}. The second dense AAM (SVS-AAMs-unconstrained) is built using deformation fields obtained by replacing the correspondence bases by identity bases so that term \ref{eq:lowrank} in the proposed variational shape flow energy becomes: 
\begin{equation}
\beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\bm{v}_n(\bx)\|^2 \ud \bx \label{}
\end{equation}
with the rest of terms (\ref{eq:costfunc} and \ref{eq:TVterm}) remaining the same. Note that this corresponds to a degenerate case where shape flow becomes equivalent to classical frame-to-frame optical flow using TV regularization on the flow $\bm{u}_n$ \cite{Garg:2013hu}.

As before, both dense AAM are build using the 811 training images of LFPW database and tested on the 224 testing images of the same database. The cumulative error distributions graph for this experiment is shown in Figure~\ref{fig:basis_ced} and the fitting statistics reported in Table~\ref{tab:basis_stats}. 
Both Figure~\ref{fig:basis_ced} and Table~\ref{tab:basis_stats} show that the incorporation of the correspondence bases leads to considerably better performant dense AAMs. This large difference in performance fully justifies the incorporation of these bases to our shape flow algorithm.

@@

In the following set of experiments we use the proposed methodology to build dense AAMs of 2 different object classes (faces and ears) and quantitatively assess their respective fitting accuracy when fitted to novel in-the-wild images of the previous objects.

### 2015-07-17 contents/implementation.tex
Final Rebuttal

\begin{figure*}[t!]
    \centering
        \includegraphics[width=\textwidth]{resources/architecture}
    \caption{Proposed pipeline.}
    \label{fig:archi}
\end{figure*}

@@

The proposed pipeline is depicted in Figure~\ref{fig:archi}. It takes as input a set of training images of a particular object class, along with (possibly inconsistent) point and/or curve line annotations.
As a first step, the shapes that correspond to the training data are consistently represented using multichannel SVS.
The next step is the application of ICP to achieve an initial simple alignment of the SVS images.
The similarity-aligned SVS images are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all shapes of the training set. In order for the flow estimation to be accurate and exploit the correlation across the different training shapes, it utilises a correspondence basis that is built on sparse correspondences given by Nonrigid ICP alignment of the annotated data. Finally, the dense correspondences that are yielded by the optical flow serve as automatic dense landmarks and are used to define a novel dense version of Active Appearance Models. The following sections discuss the different stages of the proposed pipeline in further detail.

@@

In order to fully capture the variability among most deformable object shapes annotations, we use a representation based on Support Vector Shapes (SVS) \cite{Nguyen2013}. An SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with  Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a) The representation is completely general, \eg it can be applied to sparse landmark points, curves lines or a combination of the two and (b) It fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers.

In practice, we assume that all the training images correspond to the same object category and contain a set of inconsistent points and curve line annotations. All annotations are densely sampled to yield a set of annotated landmarks per image, with this set being different for every training image. Initially, negative points get randomly sampled around sparse landmarks while annotated landmarks are positive points. Since the number of positive points is far smaller that the number of negative ones, landmarks are assigned considerably larger weights so that $N_p \times W_p=N_n \times W_n$ where $N_p, N_n$ are number of positive/negative samples and $W_p, W_n$ are their corresponding weights.

SVMs with RBF kernel function map any numbers of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on 2D space represents the actual shape of the object. Note that the decision function for SVMs can be mathematically expressed as:
\begin{equation} \label{eq:decisionfunc}
    d(\mathbf{x})=\sum_i\alpha_i \, k(\mathbf{x}_i^*,\mathbf{x})
\end{equation}
where $\mathbf{x}_i^*$ are support vectors and \mbox{$k(\mathbf{x}_i^*, \mathbf{x}) = exp(-\lambda \|\mathbf{x}_i^* -\mathbf{x}\|^2)$} is the RBF kernel .

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(\bx) \,\, ,
\end{equation}
where $\bm{v_i}(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v_i}(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v_i}(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v_i}(\bm{x})$. The above representation of shape flow constrains the correspondence vectors to lie on a subspace and therefore acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/6}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/7}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/8}
    \end{subfigure}
    \caption{Deformation field built}
    \label{fig:deformationfield}
\end{figure}

@@

\subsubsection{Variational Shape Flow Estimation}%{Multi-image Subspace Flow}

Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training SVS images $\bm{d_n}(\bx)$ by minimizing the following energy:
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^F \|\bm{d_n}(\bx+\bm{u}_n(x))-\bm{d_0}(x)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v_i}(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$, which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v_i}(\bx)$ that are spatially regularised and enforce a low-rank prior. We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v_i}(\bx)$. The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

The \textbf{first term} of the above energy (term \eqref{eq:costfunc}) is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d_0}(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d_n}(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy (term \eqref{eq:lowrank}) penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy (term \eqref{eq:TVterm}) corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $\bm{v_i}(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.

We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu}. In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with the shape flow result of the TPS interpolation described in Sec.~\ref{sec:trabasis}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using an alternating optimization w.r.t.~$\bm{v_i}(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
$\bm{v_i}(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

Figure~\ref{fig:deformationfield} demonstrates some examples of the deformation fields derived by the estimated shape flow using the aforementioned method. These results correspond to some of the training shapes, in the case of a face dataset.

@@

Applying PCA on deformations trains dense deformable shape model:
\begin{equation*}
    \bm{s_p}=\bm{\bar{s}} + \bm{U}_s\bm{p}
\end{equation*}
where $\bm{s_p}$ is deformed shape instance. $\bm{\bar{s}}$ is mean shape and $\bm{U}_s\bm{p}$ are eigenvectors with corresponding parameters $p$. Figure~\ref{fig:models} shows an instance of deformed shape and appearance model.
\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.22\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_dAAM/of_shape}
    \end{subfigure}
  	\hfill
    \begin{subfigure}[b]{0.22\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_dAAM/of_app}
    \end{subfigure}
    \caption{Dense Deformable Model}
    \label{fig:models}
\end{figure}

@@

The deformation fields obtained in section \ref{sec:shapeflow} can be effectively used to learn a naturally dense version of Active Appearance Models \cite{Cootes2001, Matthews2004}. Because all deformation fields contain the same number of pixels (the same as on the reference frame) the spatial positions $\mathbf{x}_i=(x_i, y_i)$ of these pixels can be treated as point landmarks and the deformation field themselves as dense annotations of the object's shape. Consequently, building a dense shape model reduces to normalizing these dense annotations with respect to a global similarity transform (typically using Procrustes Analysis) and applying Principal Component Analysis (PCA). The resulting shape model can be mathematically expressed as:
\begin{equation}
    \begin{aligned}
        \mathbf{s} & = \mathbf{\bar{s}} + \mathbf{S} \mathbf{p}
    \end{aligned}
    \label{eq:shape_model}
\end{equation}
where $\mathbf{\bar{s}}$ is mean shape and $\mathbf{S}$ and $\mathbf{p}$ are the shape bases and parameters respectively.

On the other hand, the texture model associated to the previous set of densified shapes can be easily obtained by warping all image onto the reference frame, making explicit use of the one-to-one correspondence between pixels on the reference frame and on the deformation fields. Once the images have been warped, the texture model is obtained by applying PCA on the warped images. The texture model is mathematically defined by the following expression:
\begin{equation}
    \begin{aligned}
        \mathbf{a} = \mathbf{\bar{a}} + \mathbf{A} \mathbf{c}
    \end{aligned}
	\label{eq:tex_model}
\end{equation}
where $\mathbf{a}$ is the mean texture, and $\mathbf{A}$ and $\mathbf{c}$ are the texture bases and parameters respectively.

Note that, both shape and texture models are directly related through the existent one-to-one correspondence between dense landmarks and pixels on the reference
frame. Therefore, in clear contrast to classic AAMs formulations \cite{Cootes2001, Matthews2004}, no motion model (piece-wise affine, thin-plates splines \cite{Bookstein1989}) is required to relate the previous shape and texture models in our dense formulation. As noted by other authors \cite{Amberg2009, Tzimiropoulos2014}, this has the advantage that all efficient compositional gradient descent algorithms for fitting AAMs \cite{Papandreou2008, Matthews2004, Amberg2009, Tzimiropoulos2013, Alabort2014} become exact under our formulation.

>>>>

\begin{figure*}[t!]
    \centering
        \includegraphics[width=\textwidth]{resources/architecture2}
    \caption{Schematic description of the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:archi}
\end{figure*}

@@

The proposed pipeline is depicted in Figure~\ref{fig:archi}. It takes as input a set of training images of a particular object class, along with inconsistent point and/or curve line shape annotations.
As a first step, the all shape annotations are consistently represented using multichannel SVS.
The next step is the application of ICP to achieve an initial simple alignment of the SVS images.
The multichannel similarity-aligned SVS images are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all the shapes in the training set. In order for the flow estimation to be accurate and exploit the existent correlation among all training shapes, it utilises a correspondence bases built from sparse correspondences obtained by applying Nonrigid ICP on the annotated data. Finally, the dense correspondences yield by optical flow serve as automatically generated dense landmark annotations and are used to define a novel dense version of Active Appearance Models. The upcoming sections discuss each of the previous stages in further detail.

@@

In order to fully capture the variability among most deformable object shapes annotations, we use a representation based on Support Vector Shapes (SVS) \cite{Nguyen2013}. A SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a) the representation is completely general, \eg it can be applied to sparse landmark points, curves lines or a combination of the two and (b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.

In practice, we assume that all the training images correspond to the same object category and contain a set of inconsistent points and curve line annotations. All annotations are densely sampled to yield a set of annotated landmarks per image, with this set being different for every training image. Initially, negative points get randomly sampled around sparse landmarks while annotated landmarks are positive points. Since the number of positive points is far smaller that the number of negative ones, landmarks are assigned considerably larger weights so that $N_p \times W_p=N_n \times W_n$ where $N_p, N_n$ are the number of positive/negative samples and $W_p, W_n$ are their corresponding weights.

SVMs with RBF kernel function map any number of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on 2D space represents the actual shape of the object. Note that the decision function for SVMs can be mathematically expressed as:
\begin{equation} \label{eq:decisionfunc}
    d(\bm{x})=\sum_i\alpha_i \, k(\bm{x}_i^*,\bm{x})
\end{equation}
where $\bm{x}_i^*$ are the so-called support vectors and \mbox{$k(\bm{x}, \bm{y}) = \exp\left(-\frac{\|\bm{x} -\bm{y}\|^2}{2 \sigma^2}\right)$} is the RBF kernel .

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/6}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/7}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/8}
    \end{subfigure}
    \caption{Exemplar deformation fields obtained using the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:deformationfield}
\end{figure}

@@

\subsubsection{Variational Shape Flow Estimation}%{Multi-image Subspace Flow}
\label{varshapeflow}

Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training SVS images $\bm{d}_n(\bx)$ by minimizing the following energy:
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^F \|\bm{d}_n(\bx+\bm{u}_n(x))-\bm{d}_0(x)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior. We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$. The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

The \textbf{first term} of the above energy \eqref{eq:costfunc} is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d}_0(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d}_n(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy \eqref{eq:lowrank} penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy \eqref{eq:TVterm} corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $\bm{v}_i(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.

We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu}. In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with the shape flow result of the TPS interpolation described in Sec.~\ref{sec:trabasis}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

Figure~\ref{fig:deformationfield} shows some examples of deformation fields derived from the estimated shape flow computed by the aforementioned method. These results correspond to exemplar training shapes in the case of a face dataset.

@@

% Applying PCA on deformations trains dense deformable shape model:
% \begin{equation*}
%     \bm{s_p}=\bm{\bar{s}} + \bm{U}_s\bm{p}
% \end{equation*}
% where $\bm{s_p}$ is deformed shape instance. $\bm{\bar{s}}$ is mean shape and $\bm{U}_s\bm{p}$ are eigenvectors with corresponding parameters $p$. Figure~\ref{fig:models} shows an instance of deformed shape and appearance model.
% \begin{figure}[h!]
%     \centering
%     \begin{subfigure}[b]{0.22\textwidth}
%             \includegraphics[width=\textwidth]{resources/Fig_dAAM/of_shape}
%     \end{subfigure}
%   	\hfill
%     \begin{subfigure}[b]{0.22\textwidth}
%             \includegraphics[width=\textwidth]{resources/Fig_dAAM/of_app}
%     \end{subfigure}
%     \caption{Dense Deformable Model}
%     \label{fig:models}
% \end{figure}

@@

The deformation fields obtained from section \ref{sec:shapeflow} can be effectively used to learn a naturally dense version of Active Appearance Models \cite{Cootes2001, Matthews2004}. Because all deformation fields contain the same number of pixels (the same as on the reference frame) the spatial positions $\bm{x}_i=(x_i, y_i)$ of these pixels can be treated as point landmarks and the deformation field themselves as dense annotations of the object's shape. Consequently, building a dense shape model reduces to normalizing these dense annotations with respect to a global similarity transform (typically using Procrustes Analysis) and applying Principal Component Analysis (PCA). The resulting shape model can be mathematically expressed as:
\begin{equation}
    \begin{aligned}
        \bm{s} & = \bm{\bar{s}} + \bm{S} \bm{p}
    \end{aligned}
    \label{eq:shape_model}
\end{equation}
where $\bm{\bar{s}}$ is mean shape and $\bm{S}$ and $\bm{p}$ are the shape bases and shape parameters respectively.

On the other hand, the texture model associated to the previous set of densified shapes can be easily obtained by warping all image onto the reference frame, making explicit use of the one-to-one correspondence between pixels on the reference frame and on the deformation fields. Once the images have been warped, the texture model is obtained by applying PCA on the warped images. The texture model is mathematically defined by the following expression:
\begin{equation}
    \begin{aligned}
        \bm{t} = \bm{\bar{t}} + \bm{T} \bm{c}
    \end{aligned}
	\label{eq:tex_model}
\end{equation}
where $\bm{t}$ is the mean texture, and $\bm{T}$ and $\bm{c}$ are the texture bases and texture parameters respectively.

Note that, both shape and texture models are directly related through the existent one-to-one correspondence between dense landmarks and pixels on the reference
frame. Therefore, in clear contrast to classic AAMs formulations \cite{Cootes2001, Matthews2004}, in our dense formulation no motion model (piece-wise affine, thin-plates splines \cite{Bookstein1989}) is required to relate the previous shape and texture models . As noted by other authors \cite{Amberg2009, Tzimiropoulos2014}, this has the advantage that all efficient compositional gradient descent algorithms for fitting AAMs \cite{Papandreou2008, Matthews2004, Amberg2009, Tzimiropoulos2013, Alabort2014} become exact under our formulation.

### 2015-07-17 contents/introduction.tex
Updates from ShareLaTeX
Final Rebuttal

To illustrate how time consuming careful annotation of the human ear is we lay down our own experience. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore,  training and fatigue largely influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the tasks of object alignment has recently attracted some attention \cite{kokkinos2007unsupervised, jojic2006escaping, jiang2009learning, liu2009simultaneous, baker2004automatic, cox2008least, cootes2004groupwise, frey2003learning}. Nevertheless, the method of \cite{antonakos2014automatic} requires a consistent set of shapes annotated with landmarks to perform deformable face congealing. Finally, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed the limited number of methods that have been applied for the task cannot be applied for arbitrary collection of images collected "in-the-wild".

In this paper, we provide a solution for annotating an object with regards to its deformations that requires considerable less effort and in the same time can be used to define statistical deformations for objects without consistent set of landmarks. That is, we argue that it is better to annotate an object with regards to a set of continues lines that describe the shape of the object. An example is provided in Figure \ref{fig:intro} where the shape of the facial parts is annotated with continuous lines instead of a number of discrete landmarks. Similarly, the shape annotations for the ear, a bottle and human body is shown in Figure \ref{fig:intro} Annotating facial images by following this procedure takes 42, 38 and 43 secs per image for faces, ears and bottles correspondingly. Furthermore, these shapes can be automatically produced by recent methods that perform discriminative segmentation of objects.

>>>>

To illustrate how time consuming careful annotation of the human ear is we lay down our own experience. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore,  training and fatigue largely influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the tasks of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, cox2008least, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed the limited number of methods that have been applied to the task cannot be directly applied to arbitrary collection of images collected "in-the-wild". On the other hand, the method of \cite{antonakos2014automatic}, which can deal with "in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

In this paper, we provide a solution for annotating an object with regards to its deformations that requires considerable less effort and in the same time can be used to define statistical deformations for objects without consistent set of landmarks. That is, we argue that it is better to annotate an object with regards to a set of continuous lines that describe the shape of the object. An example is provided in Figure \ref{fig:intro} where the shape of the facial parts is annotated with continuous lines instead of a number of discrete landmarks. Similarly, the shape annotations for the ear, a bottle and human body is shown in Figure \ref{fig:intro} Annotating facial images by following this procedure takes 42, 38 and 43 secs per image for faces, ears and bottles correspondingly. Furthermore, these shapes can be automatically produced by recent methods that perform discriminative segmentation of objects.
